---
title: "I prefer not to say"
#header-includes: |
#  \AtBeginDocument{\title[I prefer not to say]{I prefer not to say}}
date: last-modified
author: Dianyi Yang
format:
  beamer:
    theme: Madrid
link-citations: true
csl: extra/apa.csl
project:
  execute-dir: "D:/OneDrive - London School of Economics/Desktop/lse assignments/GV482/prefernottosay"
---

# Inspiration

Governments/organisations often want to give preferential treatment to disadvantaged groups. However, they rely on self-reported data for identification.

-   Even if we eliminate the the possibility of misreporting for some easily verifiable characteristics

    -   e.g. race, poverty

-   Imperfect information arises from refusal to report one's true identity

    -   "I prefer not to say"

# Model Setting (1/2)

-   Players:

    -   Government (G)

    -   Disadvantaged group ($i\in D$)

    -   Advantaged group ($i\in A$)

-   Strategies:

    -   G: level of support given to applicant $i$: $v_{i}\in[0,1]$

    -   D: report that they belong to the disadvantaged group or choose not to say

        -   $s_{i,D}=\{ D, N \}$

    -   A: report that they belong to the advantaged group or choose not to say

        -   $s_{i,A}=\{ A,N \}$

# Model Setting (2/2)

-   Payoffs:

    -   G: $U_{G}=-(v_{i}-I(i\in D))^2$

    -   D: $U_{i,D}(s=D)=\mu_{D}+\epsilon_{i,D}$

    -   A: $U_{i,A}(s=A)=\mu_{A}+\epsilon_{i,A}$

    -   $U_{i,D}(s=N)=U_{i,A}(s=N)=U(s=N)$

        -   unrealistic but good for plotting and does not affect conclusion

    -   $\mu_D>\mu_A$ due to preferential treatment

    -   $\epsilon$s are normally distributed for plotting (or can really be of any distribution)

-   Common knowledge:

    -   The utility functions

    -   The distribution of $\epsilon$s

-   Special assumption

    -   Equal sizes of $D$ and $A$ (for simplicity)

# Best strategies

For government ($G$):

$$
v_{i}=E(i\in D| s_{i})
$$

For $D$ and $A$:

-   Choose to reveal their true identity when

    $$
    U_{i,D}(s=D)>U(s=N)~~~~U_{i,A}(s=A)>U(s=N)
    $$

-   "I prefer not to say" otherwise

# Solve for $U(s=N)$ (baseline)

Parameters:

-   $\mu_{D}=1,~\mu_{A}=0$, $Var(\epsilon_{D})=Var(\epsilon_{A})=1$

-   $U(s=N)=\frac{\text{Blue Shaded Area}}{\text{Total Shaded Area}}\approx0.278$

Outcome:

-   61% of Advantaged Group refuse to self-identify

-   24% of the Disadvantaged Group refuse to self-identify

```{r, fig.width=8, fig.height=3}
library(tidyverse)
x <- seq(-1,5,0.01)
ya <- dnorm(x, 0, 1)
yd <- dnorm(x, 1, 1)
u=0.278
p1 <- data.frame(x=rep(x,2), y=c(ya, yd), group=c(rep("Advantaged",601), rep("Disadvantaged",601)))

ggplot(p1, aes(x=x, y=y, color=group)) + facet_grid(rows = vars(group)) +
  geom_line()+geom_vline(xintercept = u, linetype=2) + theme_bw() +
  geom_ribbon(data=subset(p1,x<=u),aes(x=x,ymax=y, fill=group),ymin=0,alpha=0.3) +
  theme(legend.position = "none") +
  labs(x="U(s=A or G)", y="Probability Density")

```

# What if there is more social stigma (lower $\mu_{D}$)?

Parameters:

-   $\mu_{D}=0.5,~\mu_{A}=0$, $Var(\epsilon_{D})=Var(\epsilon_{A})=1$

-   $U(s=N)=\frac{\text{Blue Shaded Area}}{\text{Total Shaded Area}}\approx0.413$

Outcome:

-   66%(↑) of Advantaged Group refuse to self-identify

-   47%(↑) of the Disadvantaged Group refuse to self-identify

```{r, fig.width=8, fig.height=3}
x <- seq(-1,5,0.01)
ya <- dnorm(x, 0, 1)
yd <- dnorm(x, 0.5, 1)
u=0.413
p2 <- data.frame(x=rep(x,2), y=c(ya, yd), group=c(rep("Advantaged",601), rep("Disadvantaged",601)))

ggplot(p2, aes(x=x, y=y, color=group)) + facet_grid(rows = vars(group)) +
  geom_line()+geom_vline(xintercept = u, linetype=2) + theme_bw() +
  geom_ribbon(data=subset(p2,x<=u),aes(x=x,ymax=y, fill=group),ymin=0,alpha=0.3) +
  theme(legend.position = "none") +
  labs(x="U(s=A or G)", y="Probability Density")

```

# What if the advantaged group is less cheeky (higher $\mu_{A}$)?

Parameters:

-   $\mu_{D}=1,~\mu_{A}=0.5$, $Var(\epsilon_{D})=Var(\epsilon_{A})=1$

-   $U(s=N)=\frac{\text{Blue Shaded Area}}{\text{Total Shaded Area}}\approx0.371$

Outcome:

-   45%(↓) of Advantaged Group refuse to self-identify

-   26%(↑) of the Disadvantaged Group refuse to self-identify

```{r, fig.width=8, fig.height=3}
x <- seq(-1,5,0.01)
ya <- dnorm(x, 0.5, 1)
yd <- dnorm(x, 1, 1)
u=0.371
p3 <- data.frame(x=rep(x,2), y=c(ya, yd), group=c(rep("Advantaged",601), rep("Disadvantaged",601)))

ggplot(p3, aes(x=x, y=y, color=group)) + facet_grid(rows = vars(group)) +
  geom_line()+geom_vline(xintercept = u, linetype=2) + theme_bw() +
  geom_ribbon(data=subset(p3,x<=u),aes(x=x,ymax=y, fill=group),ymin=0,alpha=0.3) +
  theme(legend.position = "none") +
  labs(x="U(s=A or G)", y="Probability Density")

```

# What if people only care about the preferential treatment *per se* and not about reporting?

Smaller variances:

-   $\mu_{D}=1,~\mu_{A}=0$, $Var(\epsilon_{D})=Var(\epsilon_{A})=0.5$

-   $U(s=N)=\frac{\text{Blue Shaded Area}}{\text{Total Shaded Area}}\approx0.0506$

Outcome:

-   54%(↓) of Advantaged Group refuse to self-identify

-   2.9%(↓↓) of the Disadvantaged Group refuse to self-identify

```{r, fig.width=8, fig.height=2.5}
x <- seq(-1,5,0.01)
ya <- dnorm(x, 0, .5)
yd <- dnorm(x, 1, .5)
u=0.0506
p4 <- data.frame(x=rep(x,2), y=c(ya, yd), group=c(rep("Advantaged",601), rep("Disadvantaged",601)))

ggplot(p4, aes(x=x, y=y, color=group)) + facet_grid(rows = vars(group)) +
  geom_line()+geom_vline(xintercept = u, linetype=2) + theme_bw() +
  geom_ribbon(data=subset(p4,x<=u),aes(x=x,ymax=y, fill=group),ymin=0,alpha=0.3) +
  theme(legend.position = "none") +
  labs(x="U(s=A or G)", y="Probability Density")

```

# Takeaways

-   Social stigma can be taken advantage of by the advantaged group.

-   The advantaged group's sympathy leads to lower self-identification of the disadvantaged group.

-   When people only care about the preferential treatment *per se*, both group self-identify less.

# Possible Extensions

-   Different sizes of the two groups

    -   when the disadvantaged group is larger relative to the advantaged group, more of the advantaged group choose not to self-identify (intuitively).

-   Different utility derived from the preferential treatment for different groups

    -   Realistically, the disadvantaged group tend to derive more utility from the preferential treatment

        -   Diminishing marginal utility

    -   Changes the numeric results but not the conclusion.

-   Different distributions

    -   Changes the numeric results but not the conclusion.

-   Misreporting (when the identity is not verifiable)

    -   More complex
